---
title: "Define quality threshold"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(DBI)
library(RMySQL)
library(tidyverse)
library(rlist)
con <- DBI::dbConnect(RMySQL::MySQL(), 
                      host = "localhost",
                      user = "root",
                      dbname = "cagablea",
                      password = "")
```

## A list of all the requests calling samples:
The function below generates a list with the requests to download the samples 
Treshholds is a variable, where the cutoff on coverage, p.val, alt_freq and others can be added:
```{r}
samples <- function(treshholds){
  samples <- list()
  req <- "SELECT * FROM `ww_ivar`,`ww_barcode_decode` WHERE ww_barcode_decode.barcode = ww_ivar.filename" 
  samples$oct <- "AND ww_ivar.filename LIKE BINARY '%210224%'" 
  samples$nov <- "AND ww_ivar.filename LIKE BINARY '%20201120%'"
  samples$dec <- "AND ww_ivar.filename LIKE BINARY '%20201223%'"
  samples$jan <- "AND ww_ivar.filename LIKE BINARY '%210127%'"
  samples$fev <- "AND ww_ivar.filename LIKE BINARY '%20210302%'"
  samples$mar <-  "AND ww_ivar.filename LIKE BINARY '%210325%'"
  samples <- lapply(samples, function(x) paste(req, treshholds, x, " "))
  return(samples)
}
treshholds <- "AND ww_ivar.total_dp > 10 AND ww_ivar.pval < 0.05 AND ww_ivar.alt_freq > 0.03"
samples1 <- samples(treshholds)

```
Count mutations not associated to any lineage and the mutations described for at least one lineage:
TRUE correspons to the mutations that are not described for any lineages
```{r}
no_lineage <- function(d_request){
  feb <- dbGetQuery(con, d_request)
  feb <- feb[,-1]
  feb <- distinct(feb, pos, alt, .keep_all = F)
  #how many mutations are not associated with any lineage?
  pango <- dbGetQuery(con, "SELECT * FROM `pango_lineages`")
  all.mut.id <-  feb %>% left_join(pango, by = c("pos"="pos", "alt"="alt"), keep = T )
  all.mut.id <- all.mut.id %>% select(pos.x, ref, alt.x, lineage )%>% distinct(pos.x, alt.x, ref, .keep_all = T) 
  #all.mut.id
  all.mut.id %>% count(is.na(lineage))
   
}
#no_lineage(samples1$mar)#test
#Positions with nonsynonymous mutations in spike region not associated with lineages (top 2000 from covidcg)
  #lin_na %>% filter (ref_aa != alt_aa) %>% filter(pos.x >= 21563) %>% filter(pos.x <= 25384) 
n_mut <- function(samples){
  n_id_mutations <- lapply(samples, no_lineage)
  #n_id_mutations #the number of id and unidentified mut, min number of reads 10
  n_mut <- cbind(n_id_mutations$oct, n_id_mutations$nov$n, n_id_mutations$dec$n, n_id_mutations$jan$n, n_id_mutations$fev$n, n_id_mutations$mar$n)
  names(n_mut) <- c("lin_not_id", "oct", "nov", "dec", "jan", "fev", "mar")
  n_mut
}  
```
total_dp > 10 AND ww_ivar.pval < 0.05 AND ww_ivar.alt_freq > 0.03
```{r}
treshholds <- "AND ww_ivar.total_dp > 10 AND ww_ivar.pval < 0.05 AND ww_ivar.alt_freq > 0.03"
samples1 <- samples(treshholds)
n_mut(samples1)

```
Minimum number of reads 100, alt_freq 0.1, pval<0.01:
```{r}
treshholds <- "AND ww_ivar.total_dp > 100 AND ww_ivar.pval < 0.01 AND ww_ivar.alt_freq > 0.1"
samples1 <- samples(treshholds)
n_mut(samples1)

```
Minimum number of reads 100, alt_freq 0.3, pval<0.01:
```{r}
treshholds <- "AND ww_ivar.total_dp > 100 AND ww_ivar.pval < 0.01 AND ww_ivar.alt_freq > 0.3"
samples1 <- samples(treshholds)
n_mut(samples1)

```
```{r}
treshholds <- "AND ww_ivar.total_dp > 100 AND ww_ivar.pval < 0.01 AND ww_ivar.alt_freq > 0.5"
samples1 <- samples(treshholds)
n_mut(samples1)

```
```{r}
treshholds <- "AND ww_ivar.total_dp > 100 AND ww_ivar.pval < 0.01 AND ww_ivar.alt_freq > 0.7"
samples1 <- samples(treshholds)
n_mut(samples1)

```
Loop through requests with different p-value, total_dp > 100 and alt_freq > 0.1
Create a list with the sequence of p-values to test:
```{r}
treshholds_p_list <- list()
q <- "AND ww_ivar.total_dp > 100 AND ww_ivar.alt_freq > 0.1 AND ww_ivar.pval < "
p_values <- c(10 %o% 10^(-3:-16))
i = 1
for (i in 1:length(p_values)){
  treshholds_p_list <- list.append(treshholds_p_list, paste(q, p_values[i]) )
  i = i +1
}

```
Create a serie of requests with different p-values for every month:
```{r}
samples1 <- samples(treshholds_p_list) 
```
n_mut_pval has 14 dataframes, that correspond to p-values in p_values vector 
1e-02 1e-03 1e-04 1e-05 1e-06 1e-07 1e-08 1e-09 1e-10 1e-11 1e-12 1e-13 1e-14 1e-15
```{r}
i = 1
n_mut_pval <- list()
for (i in 1:length(samples1[[1]])){
  n_mut_pval <- list.append(n_mut_pval, n_mut(lapply(samples1, `[[`, i)))
  i = i + 1
}

```

Build a function to plot the number of mutatios against log10(p-value):

```{r}
plot_mut_pval <- function (month){
pval_nmut <- rbind(n_mut_pval[[1]][,month], n_mut_pval[[2]][,month])

for (i in 3:length(n_mut_pval)) {
  pval_nmut <- rbind(pval_nmut, n_mut_pval[[i]][,month] )
  pval_nmut
  i = i +1
}
pval_nmut<- as.data.frame(cbind(pval_nmut, p_values))
names(pval_nmut) <- c("identified", "not_id", "p_values")
print(pval_nmut)
ggplot(pval_nmut)+
  geom_point(aes(y = identified, x = log10(p_values), col = "red"))+
  geom_point(aes(y = not_id, x = log10(p_values), col = "blue"))+
  scale_y_continuous(name="Number of mutations")+
  theme(legend.position="none")
  
#plot(y = pval_nmut$identified, x = log10(pval_nmut$p_values))
#plot(y = pval_nmut$not_id, x = log10(pval_nmut$p_values)) 
}

```

## plot the number of mutatios against log10(p-value)
plot for october:
```{r}
month = "oct"
plot_mut_pval(month)
```

plot for november:
```{r}
month = "nov"
plot_mut_pval(month)
```

plot for december:
```{r}
month = "dec"
plot_mut_pval(month)
```

plot for january:
```{r}
month = "jan"
plot_mut_pval(month)
```

plot for february:
```{r}
month = "fev"
plot_mut_pval(month)
```

plot for march:
```{r}
month = "mar"
plot_mut_pval(month)
```

## Frequency values

Fix p-value 0.01, total_dp > 100. Create a serie of requests with alt_freq:
```{r}
treshholds_fr_list <- list()
q <- "AND ww_ivar.total_dp > 100  AND ww_ivar.pval < 0.01 AND ww_ivar.alt_freq > "
fr_values <- seq(0.03, 0.5, by = 0.025 )
i = 1
for (i in 1:length(fr_values)){
  treshholds_fr_list <- list.append(treshholds_fr_list, paste(q, fr_values[i]) )
  i = i +1
}

```

Create a serie of requests with different frequency values for every month:
```{r}
samples2 <- samples(treshholds_fr_list) 
#samples2
```
n_mut_pval has 19 dataframes, that correspond to p-values in p_values vector 
1e-02 1e-03 1e-04 1e-05 1e-06 1e-07 1e-08 1e-09 1e-10 1e-11 1e-12 1e-13 1e-14 1e-15
this chunk takes long to be executed
```{r}
i = 1
n_mut_frval <- list()
for (i in 1:length(samples2[[1]])){
  n_mut_frval <- list.append(n_mut_frval, n_mut(lapply(samples2, `[[`, i)))
  i = i + 1
}

```
Build a function to plot the number of mutatios against frequecies:

```{r}
plot_mut_freq <- function (month){
nmut <- rbind(n_mut_frval[[1]][,month], n_mut_frval[[2]][,month])

for (i in 3:length(n_mut_frval)) {
  nmut <- rbind(nmut, n_mut_frval[[i]][,month] )
  nmut
  i = i +1
}
nmut<- as.data.frame(cbind(nmut, fr_values))
names(nmut) <- c("identified", "not_id", "frequencies")
print(nmut)
ggplot(nmut)+
  geom_point(aes(y = identified, x = frequencies, col = "red"))+
  geom_point(aes(y = not_id, x = frequencies, col = "blue"))+
  scale_y_continuous(name="Number of mutations")+
  theme(legend.position="none")
  
#plot(y = pval_nmut$identified, x = log10(pval_nmut$p_values))
#plot(y = pval_nmut$not_id, x = log10(pval_nmut$p_values)) 
}

```

plot the number of mutations in october agains diff. frequencies:
Red color = Not identified mut
Blue color = Identified
```{r}
plot_mut_freq(month = "oct")
```

November:
```{r}
plot_mut_freq(month = "nov")
```

December:
```{r}
plot_mut_freq(month = "dec")
```

January:
```{r}
plot_mut_freq(month = "jan")
```

February:
```{r}
plot_mut_freq(month = "fev")
```

March:
```{r}
plot_mut_freq(month = "mar")
```